---
title: "SICSS Day 1: NLP"
subtitle: "Exercise 1 - Counting Words"
output:
  html_document:
    code_folding: hide
date: "`r format(Sys.time(), '%d %B, %Y')`"

author: "Alejandro Hermida Carrillo"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 5)
```

## Setup!
### Load the Necessary Packages
```{r libs, include=FALSE}
## Load the libraries

# Expression that loads a package by name
req <- substitute(require(x, character.only = TRUE))
# List of libraries
libs <- c("data.table", "ggplot2","tidyverse","skimr")
# Attempt to load each package; if not installed, install it and try again
sapply(libs, function(x) eval(req) || {install.packages(x, dependencies = TRUE); eval(req)})

```

### Set Working Directory and Load Helpers

```{r wd}
# Set working directory    -----------------------------------------------------
path <- dirname(rstudioapi::getActiveDocumentContext()$path) # This is a relative path. Set alternative path if necessary
setwd(path)

#Load psychological dictionaries on agency and communality (from Lawson et al., 2022)
source("helpers/SICSS_day1_helpers_dicts.R")

```

### Let's load some cool data. **Political speeches** from Trump and Biden (2020, CITE)

```{r data}

# Load it
speeches <- readRDS("data/Biden_Trump_2020_speeches.rds")

```

## **Excercise 1.1:** who has a more "communal" linguistic style? and who a more "agentic" one?

### First, let's inspect the psych dictionaries and the data. Is the data already in a state in which it can be analyzed?

#### The psych. dictionaries: What's in them?
```{r dicts-view}
str(psych_dicts)
```
##### Do we need all of them?

#### Now the data: What's in it?
```{r data-view}
skim(speeches)
str(speeches)
```

#### Visual inspection shows that the speeches are transcribed inconsistently in uppercase / lowercase. Needs a litte bit more cleaning.

```{r}
speeches_clean <- speeches %>%
  mutate(
    #lowercase is the most standard pre-processing step
    text = tolower(raw_text),
    #Let's just use the last name of each candidate as identifier
    candidate_last = str_to_upper(word(candidate, -1)),
  ) %>%
  select(candidate = candidate_last, speech_ID, date, text) %>%
  as.data.table()

```

#### Now, let's compare the language of both candidates, deliver nice plots, statistical tests, and conclusions

```{r}

```

## **Excercise 1.2:** Do the expressed communality and agency of candidates change after crisis events? 

### Compare both candidates, deliver nice plots, statistical tests, and conclusions

#### *Hint:* Is the data ready to be analyzed? We probably need to add some information

```{r}

```

## **Excercise 1.3** (if there is time): who has a richer lexicon?

```{r}

```

